World Heritage Sites on Wikipedia: Inscription Controversies in a Context of Constrained Agency for Cultural Heritage Activism

### 2020-05-30 analysis/paper/paper.Rmd
starting discussion

Only one of the controversial sites discussed above is in the group of articles where the talk page is longer than the article itself, and none of the site on the UNESCO heritage in danger list are in this group. The talk page for Mapungubwe is lengthy with expressions of concern about the accuracy of information in the article. In particular, editors were concerned about the depiction of Indigenous people, especially the degree to which early European colonists were aware of Indigenous communities, and the correct cultural affiliation of the Indigenous groups that originally occupied the site. Although not notable for the length of its talk page, the discussions among editors of the article about Preah Vihear Temple

@@

Figure \@ref(fig:wh-wiki-map) shows that most countries have between 2-3 CS-WHL on Wikipedia, with Germany, India, China, France and Spain being the countries with the most articles about their CS-WHL.

@@

>>>>

Only one of the controversial sites discussed above is in the group of articles where the talk page is longer than the article itself, and none of the site on the UNESCO heritage in danger list are in this group. The talk page for Mapungubwe is lengthy with expressions of concern about the accuracy of information in the article. In particular, editors were concerned about the depiction of Indigenous people, especially the degree to which early European colonists were aware of Indigenous communities, and the correct cultural affiliation of the Indigenous groups that originally occupied the site. Although not notable for the length of its talk page, the discussions among editors of the article about Preah Vihear Temple include a strongly worded disagreement among seven editors about the legitimacy of the ownership claims by Thailand and Cambodia of the territory that includes the temple. The debate among Wikipedia editors mirrors very closely the political tensions surrounding the nomination of Preah Vihear. The talk page also has a much longer argument, mostly between two editors, about whether or not the Thai name of the site should be included in the text of the article. This debate also recapitiulates issues of which country has the stronger claim for ownership of the temple. Amid accusations of nationalism among the editors, one of them asks "Can we put aside politics and be more collaborative on the encyclopedia project?" and appeals to Wikipedia policies 'Third opinion' and 'Dispute resolution' to diffuse the tensions and refocus attention on the common goal. Beyond these two pages we found that the majority of content on CS-WHL article talk pages is about highly local topics, mostly article administration, editing processes, and quality control.

@@

```{r wh-wiki-circle, fig.cap="Flow of edits by location on Wikipedia for articles about Cultural Sites on the UNESCO World Heritage List. The arrows indicate the country where the editor is location (arrow nock or start) and the location of the CS-WHL site that they are editing the article of (arrow point or end). Editor locations were determined by geo-locating the IP addresses associated with anonymous edits. Inset shows the distribution of proportions of edits per article that are anonymous. Non-anonymous edits are identitifed by Wikipedia usernames, rather than IP addresses, and cannot be geo-located.", fig.width=15}
source(here::here("analysis/paper/007-Wikipedia-WHL-spatial-circle-plot.R"))
knitr::include_graphics(here::here("analysis/figures/wh_countries_circle_chord_fig.png"))
```

Spatial patterns in the coverage of CS-WHL sites on Wikipedia closely resemble the physical distribution of sites. Figure \@ref(fig:wh-wiki-map) shows that most countries have between 2-3 CS-WHL on Wikipedia, with Germany, India, China, France and Spain being the countries with the most articles about their CS-WHL. This pattern is consistent with @graham2014uneven's observations that places in the Global North tend to be over-represented on Wikipedia. More surprising is the distribution of editor locations shown in figure \@ref(fig:wh-wiki-circle), which we determined by geo-locating the IP addresses attached to each anonymous edit on all CS-WHL articles. This figure shows the flow of edits, with the arrows starting at the country where the editor was located when they made their edit, and ending at the country where the CS-WHL site that they are editing is located. The inset plot on figure \@ref(fig:wh-wiki-circle) shows that this visualisation only accounts for a relatively small proportion of all edits to CS-WHL articles. Around 20% of edits of most CS-WHL articles are anonymous, a higher proportion than our random sample. Nevertheless, there are `r prettyNum(nrow(revision_history_page_details_ip_addresses_df), big.mark=",")` anonymous edits, and so this sample likely contains a non-random signal. The most striking detail in figure \@ref(fig:wh-wiki-circle) is the large proportion of edits that originate in the United States, and that the vast majority of these are edits of articles about CS-WHL sites located elsewhere in the world. The country with the next largest proportion, the United Kingdom, is less than half of the United States, and a much greater proportion of edits originating in the UK are on articles about CS-WHL sites located in the UK, compared to the US. The country with the next highest proportion of editors is India, and nearly half of those edits are on CS-WHL sites located India. The majority of countries in this chord diagram indicate that articles about CS-WHL sites in those countries receive edits from editors located elsewhere, mostly the US and other English-speaking countries.

```{r wh-wiki-time1, fig.cap="...", fig.width=15}
source(here::here("analysis/paper/008-Wikipedia-WHL-time-series.R"))
knitr::include_graphics(here::here("analysis/figures/wh_wikipedia_time_inscription.png"))
```

```{r wh-wiki-time2, fig.cap="...", fig.width=15}
knitr::include_graphics(here::here("analysis/figures/wh_wikipedia_time_series_facet.png"))
```

Time series analysis of edits on articles about CS-WHL sites can help us identify relationships between activity on Wikipedia articles and external events related to the sites. Figure \@ref(fig:wh-wiki-time1) shows that most articles about CS-WHL sites are much older than articles in our random sample, with most CS-WHL articles created 5-10 years after Wikipedia first appeared in 2001. This rapid addition of articles early in the life of Wikipedia supports the observation from the in-bound Wikilinks data that articles about CS-WHL sites are more central to the encyclopedia and considered worthy of inclusion than many other types of articles. The distribution of years between the date of a site's inscription on the WHL and the date of the creation of that site's Wikipedia article has a distinctive trimodal shape. The first two modes reflect the relatively large number of sites inscribed on the WHL in 1983 and the early 1990s. The third mode at the zero point on the horizontal axis corresponds to articles that were created in or around the year the CS-WHL site was inscribed on the list. These are sites inscribed after the creation of Wikipedia in 2001. Only a relatively small proportion of CS-WHL sites had Wikipedia articles created about them before they were inscribed on the WHL, because the majority of sites were inscribed before Wikipedia was created. The time series of editing activity at individual sites are highly variable and show display complex patterns (Figure \@ref(fig:wh-wiki-time2)). Timbuktu shows spikes of activity right after attacks in mid June 2012, with descriptions of the attacks added to the text of the article. However the tallest peak in editing activity on the Timbuktu occurs earlier, in April 2012 that  concentrated on expanding . A detailed analysis of these is beyond the scope of this paper, but we can see several patterns that suggestion interesting relationships between editorial activity and WHL listing. For example, editing activity spiking sharply at the time of WHL inscription (e.g. Tyre, Masada), editorial activity spiking at the time of WHL inscription and slowly decaying to a baseline level (e.g. Carthage), editorial activity that peaks shortly after the time of WHL inscription (e.g. Troy, Masada), and many sites where editorial activity that shows no signal at all at the time of WHL inscription (Figure \@ref(fig:wh-wiki-time2)).

@@

https://en.wikipedia.org/wiki/Wikipedia:Geopolitical,_ethnic,_and_religious_conflicts_noticeboard now defunct
